---
output:
  md_document:
    variant: markdown_github
---
```{r Setup, include=FALSE}
library(gasper)
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align = "center",
  out.width = '50%',
  dpi = 150,
  comment = "#>"
)
```
# gasper

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/gasper)](http://cran.r-project.org/package=gasper)
![CRAN Downloads](http://cranlogs.r-pkg.org/badges/gasper)
![CRAN Downloads](http://cranlogs.r-pkg.org/badges/grand-total/gasper)

Graph signal processing in R.
```{r,echo=FALSE}
f <- exp(-( (rlogo$xy[,1]-1)^2/2+(rlogo$xy[,2]-4)^2/2))
plot_signal(rlogo,f,size = 3)
```

## Download and Install

Install the devtools package if you haven't already.
```{r, eval=FALSE}
install.packages("devtools")
```

To install the development package, type the following at the R command line:
```{r, eval=FALSE}
devtools::install_github("fabnavarro/gasper")
library(gasper)
```

To install the CRAN version of the package, type the following:
```{r, eval=FALSE}
install.packages("gasper")
```


To obtain the complete list of package functions, simply type 
```{r, eval=FALSE}
help(package = "gasper")
```

## Getting Started

See the [package vignette](https://fnavarro.perso.math.cnrs.fr/rpackage/gasper_vignette.pdf) or [documentation](https://fnavarro.perso.math.cnrs.fr/rpackage/gasper.pdf) for more details.
You could also build and see the vignette associated with the package using the following lines of code
```{r, eval=FALSE}
devtools::install_github("fabnavarro/gasper", build_vignettes = TRUE)
library(gasper)
```

Then, to view the vignette
```{r, eval=FALSE}
vignette("gasper_vignette")
```

For an illustration of the features of the package, you can also refer to the following repo [SGWT-SURE](https://github.com/fabnavarro/SGWT-SURE) which provides an effective generalization of the Stein Unbiased Risk Estimate (SURE) for signal denoising/regression on graphs using Spectral Graph Wavelet Transform.

## Interface to the SuiteSparse Matrix Collection

The package also provides an interface to the SuiteSparse Matrix Collection, which is a large and actively growing set of sparse matrix benchmarks gathered from a broad spectrum of applications (for details see https://sparse.tamu.edu/).

The `download_graph` function allows to download a test matrix from this collection. For example:
```{r, message=FALSE}
matrixname <- "usroads-48"
groupname <- "Gleich"
download_graph(matrixname,groupname)
attributes(`usroads-48`)
```
`usroads-48` is composed of the sparse matrix `sA` (in compressed sparse column format), coordinates `xy` (if present, in a data.frame), `dim` the number of rows, columns and numerically nonzero elements and `info` about the matrix (stored in a temporary folder as a plain text file) that can be display via `file.show(usroads-48$info)` for example or in the console:

```{r}
cat(readLines(`usroads-48`$info), sep = "\n")
```

It is also possible to plot a (planar) graph and plot signals defined on top of it. For example :
```{r,fig.show='hold'}
f <- rnorm(nrow(`usroads-48`$xy))
plot_graph(`usroads-48`, size = 0.05)
plot_signal(`usroads-48`, f, size = f/4)
```

In cases where these coordinates are not supplied, `plot_graph` employs simple spectral graph embedding to calculate some node coordinates (nodes that are connected or share structural similarities in the graph are placed close to each other in the spectral drawing). This is done using the function `spectral_coords`, which computes the spectral coordinates based on the eigenvectors associated with the smallest non-zero eigenvalues of the graph's Laplacian.
```{r, message=FALSE,fig.show='hold'}
matrixname <- "delaunay_n10"
groupname <- "DIMACS10"
download_graph(matrixname,groupname)
attributes(delaunay_n10)
plot_graph(delaunay_n10)
plot_signal(delaunay_n10,
            cos(1:nrow(delaunay_n10$sA)))
```


