---
output:
  md_document:
    variant: markdown_github
---
```{r Setup, include=FALSE}
library(gasper)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# gasper

[![Travis build status](https://travis-ci.org/fabnavarro/gasper.svg?branch=master)](https://travis-ci.org/fabnavarro/gasper)

Graph signal processing in R.
```{r,echo=FALSE}
f <- exp(-( (rlogo$xy[,1]-1)^2/2+(rlogo$xy[,2]-4)^2/2))
plot_signal(rlogo,f,size = 3)
```

## Download and Install

Install the devtools package if you haven't already.
```{r, eval=FALSE}
install.packages("devtools")
```

To install the development package, type the following at the R command line:
```{r, eval=FALSE}
devtools::install_github("fabnavarro/gasper")
library(gasper)
```

To obtain the complete list of package functions, simply type 
```{r, eval=FALSE}
help(package = "gasper")
```

## Getting Started

See the [package vignette](http://fnavarro.perso.math.cnrs.fr/rpackage/gasper_vignette.pdf) for more details.
You could also build and see the vignette associated with the package using the following lines of code
```{r, eval=FALSE}
devtools::install_github("fabnavarro/gasper", build_vignettes = TRUE)
library(gasper)
```

Then, to view the vignette
```{r, eval=FALSE}
vignette("gasper_vignette")
```

For an illustration of the features of the package, you can also refer to the following repo [SGWT-SURE](https://github.com/fabnavarro/SGWT-SURE) which provides an effective generalization of the Stein Unbiased Risk Estimate (SURE) for signal denoising/regression on graphs using Spectral Graph Wavelet Transform.

